### 2025-02-15 paper/preprint/paper.tex
interesting

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Parameters}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{corollary}{Research Question}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

In panel (c), we have used conventional training again, this time
imposing the mutability constraint on \emph{age} at test time.
Counterfactuals are valid but involve some substantial reductions in
\emph{debt} for some individuals. By comparison, counterfactual paths
are shorter on average in panel (d), where we have used counterfactual
training and protected immutable features as described in
Section~\ref{sec-constraints}. The counterfactuals are also plausible
with respect to the mutable feature. Thus, we consider the model in
panel (d) as the most explainable according to
Definition~\ref{def-explainability}.

@@

\begin{tcolorbox}[enhanced jigsaw, opacityback=0, breakable, left=2mm, coltitle=black, colbacktitle=quarto-callout-warning-color!10!white, titlerule=0mm, opacitybacktitle=0.6, colback=white, bottomtitle=1mm, toptitle=1mm, bottomrule=.15mm, arc=.35mm, colframe=quarto-callout-warning-color-frame, rightrule=.15mm, title=\textcolor{quarto-callout-warning-color}{\faExclamationTriangle}\hspace{0.5em}{Warning}, toprule=.15mm, leftrule=.75mm]

@@

\begin{tcolorbox}[enhanced jigsaw, opacityback=0, breakable, left=2mm, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, opacitybacktitle=0.6, colback=white, bottomtitle=1mm, toptitle=1mm, bottomrule=.15mm, arc=.35mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, title=\textcolor{quarto-callout-note-color}{\faInfo}\hspace{0.5em}{Note}, toprule=.15mm, leftrule=.75mm]

>>>>

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{corollary}{Research Question}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Parameters}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

In panel (c), we have used conventional training again, this time
imposing the mutability constraint on \emph{age} at test time.
Counterfactuals are valid but involve some substantial reductions in
\emph{debt} for some individuals, in particular very young applicants.
By comparison, counterfactual paths are shorter on average in panel (d),
where we have used counterfactual training and protected immutable
features as described in Section~\ref{sec-constraints}. In particular,
we observe that due to the classifier's lower sensitivity to \emph{age},
recourse recommendations with respect to \emph{debt} are much more
homogenous, in that they do not disproportionately punish younger
individuals. The counterfactuals are also plausible with respect to the
mutable feature. Thus, we consider the model in panel (d) as the most
explainable according to Definition~\ref{def-explainability}.

@@

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-warning-color-frame, toprule=.15mm, opacityback=0, left=2mm, breakable, colback=white, bottomtitle=1mm, rightrule=.15mm, titlerule=0mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-warning-color!10!white, title=\textcolor{quarto-callout-warning-color}{\faExclamationTriangle}\hspace{0.5em}{Warning}, coltitle=black, toptitle=1mm, arc=.35mm, bottomrule=.15mm]

@@

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toprule=.15mm, opacityback=0, left=2mm, breakable, colback=white, bottomtitle=1mm, rightrule=.15mm, titlerule=0mm, opacitybacktitle=0.6, colbacktitle=quarto-callout-note-color!10!white, title=\textcolor{quarto-callout-note-color}{\faInfo}\hspace{0.5em}{Note}, coltitle=black, toptitle=1mm, arc=.35mm, bottomrule=.15mm]

### 2025-02-15 paper/sections/method.qmd
interesting

In panel (c), we have used conventional training again, this time imposing the mutability constraint on *age* at test time. Counterfactuals are valid but involve some substantial reductions in *debt* for some individuals. By comparison, counterfactual paths are shorter on average in panel (d), where we have used counterfactual training and protected immutable features as described in @sec-constraints. The counterfactuals are also plausible with respect to the mutable feature. Thus, we consider the model in panel (d) as the most explainable according to @def-explainability.

>>>>

In panel (c), we have used conventional training again, this time imposing the mutability constraint on *age* at test time. Counterfactuals are valid but involve some substantial reductions in *debt* for some individuals, in particular very young applicants. By comparison, counterfactual paths are shorter on average in panel (d), where we have used counterfactual training and protected immutable features as described in @sec-constraints. In particular, we observe that due to the classifier's lower sensitivity to *age*, recourse recommendations with respect to *debt* are much more homogenous, in that they do not disproportionately punish younger individuals. The counterfactuals are also plausible with respect to the mutable feature. Thus, we consider the model in panel (d) as the most explainable according to @def-explainability.

### 2025-02-15 paper/sections/other/constraints.qmd
interesting

    # Counterfactuals:
    if !any(isnothing.(x))
        yhat = [argmax(y) for y in eachcol(model(x))]
        yhat0 = [argmax(y) for y in eachcol(model(x0))]
        if any(yhat.==2)

            # Paths:
            u = []
            v = []
            for (i,ce) in enumerate(eachcol(x))
                Δ = ce - x0[:,i]
                push!(u, Δ[1])
                push!(v, Δ[2])
            end
            Plots.quiver!(x0[1,yhat.==2], x0[2,yhat.==2], quiver=(u[yhat.==2], v[yhat.==2]), color=1)

            # End points:
            Plots.scatter!(x[1,yhat.==2], x[2,yhat.==2], label=["CE (y⁺=1)" "CE (y⁺=2)"], ms=15, shape=:star, color=yhat[yhat.==2], group=yhat[yhat.==2], mscolor=yhat0[yhat.==2])

@@

# Data:
specs = [
    ("(a)", VanillaObjective(needs_ce=true), ["both", "both"]),
    ("(b)", EnergyDifferentialObjective(lambda=[1.0,1.0,0.0]), ["both", "both"]),
    ("(c)", VanillaObjective(needs_ce=true), ["both", "none"]),
    ("(d)", EnergyDifferentialObjective(lambda=[1.0,1.0,0.0]), ["both", "none"]),
]
```

@@

    # Objective:
    generator = GenericGenerator()
    opt_state = Flux.setup(Descent(), model)
    conv = MaxIterConvergence()

    model, logs = counterfactual_training(
        obj,
        model,
        generator,
        train_set,
        opt_state;
        val_set = val_set,
        nepochs = 50,
        mutability = Symbol.(constraint),
        callback = plot_db,
        nce=50,
        convergence=conv, 
    )

>>>>

    # Counterfactuals:
    target = 2
    factual = 1
    if !any(isnothing.(x))
        yhat = [argmax(y) for y in eachcol(model(x))]
        yhat0 = [argmax(y) for y in eachcol(model(x0))]
        idx_plotted = yhat.==target .&& yhat0.==factual
        if any(idx_plotted)

            # Paths:
            u = []
            v = []
            for (i,ce) in enumerate(eachcol(x))
                Δ = ce - x0[:,i]
                push!(u, Δ[1])
                push!(v, Δ[2])
            end
            Plots.quiver!(x0[1,idx_plotted], x0[2,idx_plotted], quiver=(u[idx_plotted], v[idx_plotted]), color=factual)

            # End points:
            Plots.scatter!(x[1,idx_plotted], x[2,idx_plotted], label=["CE (y⁺=1)" "CE (y⁺=2)"], ms=15, shape=:star, color=yhat[idx_plotted], group=yhat[idx_plotted], mscolor=yhat0[idx_plotted])

@@

# Data:
specs = [
    ("(a)", VanillaObjective(needs_ce=true), ["both", "both"]),
    ("(b)", FullObjective(lambda=[1.0,1.0,0.0,1.0]), ["both", "both"]),
    ("(c)", VanillaObjective(needs_ce=true), ["both", "none"]),
    ("(d)", FullObjective(lambda=[1.0,1.0,0.0,1.0]), ["both", "none"]),
]
```

@@

    # Objective:
    generator = GenericGenerator()
    opt_state = Flux.setup(Descent(), model)
    conv = MaxIterConvergence(5)

    model, logs = counterfactual_training(
        obj,
        model,
        generator,
        train_set,
        opt_state;
        val_set = val_set,
        nepochs = 50,
        mutability = Symbol.(constraint),
        callback = plot_db,
        nce=50,
        convergence=conv, 
        verbose=3
    )
